### 2015-07-09 paper/abbreviations.tex
add new intro

% -*- root: main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-06-16 15:45:32
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-07-08 12:12:18

\newacronym{dmri}{dMRI}{diffusion MRI}
\newacronym{hcp}{HCP, \cite{essen_human_2012}}{Human Connectome Project}
\newacronym{fod}{FOD}{fiber orientation distribution}
\newacronym{wm}{WM}{white matter}
\newacronym{snr}{SNR}{signal-to-noise ratio}
\newacronym{csd}{CSD}{constrained spherical deconvolution}
\newacronym{cc0}{CC0}{Creative Commons Zero licence}
\newacronym{roc}{ROC}{receiver operating characteristic}

\makeglossaries